---
title: "RapTLR Training Guide"
subtitle: "Rapid Top-Line Report Generation with R"
author: "RapTLR Development Team"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
description: "Complete training material for the RapTLR package"
output: word_document
---


# Introduction {#intro}

## What is RapTLR?

RapTLR is an R package designed to automate and streamline the creation of Top-Line Results Reports (TLRs) in clinical trials. It eliminates manual, error-prone tasks by:

- **Automatically appending TLFs** (Tables, Listings, Figures) to Word documents with hyperlinks
- **Dynamically replacing placeholder text** with calculated values from your data
- **Creating formatted text descriptions** of variable distributions

## Why Use RapTLR?

Traditional TLR creation involves:

- Manual copy-pasting of outputs into Word documents
- Manually updating numbers when data changes
- Risk of transcription errors
- Time-consuming formatting

**RapTLR solves these problems** by connecting your R analysis directly to your Word document, ensuring accuracy and time saving.

## Installation

```{r eval=FALSE}
# Install remotes if needed
install.packages("remotes")

# Install RapTLR from GitHub
library(remotes)
remotes::install_github("johnsonandjohnson/RapTLR")
```

## Getting Started with Examples

RapTLR includes a complete built-in example to help you learn quickly:

```{r eval=FALSE}
library(RapTLR)

# Export example files to your working directory
RapTLR::import_demo() 
```

This creates a script called `demo_TLR` that you can open, run and navigate. We provided some example shells for :
- TLR template word document : `TLR_Shell.docx`
- TLR table of list of outputs : `TLF_list.xlsx`
- TLR outputs (folder `TLF_outputs`)

You can access all of them in the dedicated [GitHub repository](https://github.com/johnsonandjohnson/RapTLR/tree/main/inst/extdata) 

## Typical Workflow

```{r eval=FALSE}
# 1. Prepare a Word shell document with placeholders
# 2. Load your data and your rtf TLFs 
# 3. Transform your rtf TLFs as docx TLFs using 3C Helper
# 4. Use run_apdx() to add TLFs in the TLR shell's appendice 
# 5. Use textReplace() to update placeholders using built-in functions 
# 6. Generate final TLR document
```

## Package Dependencies

RapTLR relies on the `officer` package for Word document manipulation. Key concepts:

- Documents are R objects that can be modified
- Changes can be chained together
- Final output is written to as a docx file

Let's dive into each component!
